### 2019-07-18 mathematical-background/information-theory.tex
hamming, jaccard distances

>>>>

\subsection{Hamming Distance} \label{subsec:hamming-distance}
The \textit{Hamming Distance} is a widely-used distance measure that quantifies the similarity of strings.
\begin{definition}
	The Hamming Distance $d_H(x,y)$ between two vectors $x$ and $y$ is given by:
	\begin{equation}
		d_H(x,y) = \sum_i \Gamma(x_i, y_i)
	\end{equation}
	with $\Gamma(i,j)$ defined as:
	\begin{equation}
		\Gamma(i,j) = 
		\begin{cases}
			0 \quad i \neq j \\
			1 \quad i = j	
		\end{cases}
	\end{equation}
\end{definition}
Given strings of characters, or more in general vectors over some field, of equal length, their Hamming Distance is the number of positions where they differ.
It can be seen as the number of substitutions needed to transform one into the other.

This is a valid distance measure because:
\begin{itemize}
  \item it is non-negative: $d_H(x,y) \geq 0 \quad \forall x, y$
  \item it fulfils the identity of indiscernibles: $x = y \Rightarrow d_H(x,y)=0$
  \item it respects the \textit{Triangle Inequality}: $d_H(x,y) \leq d_h(x,z) + d_H(z,y) $
\end{itemize}

For example, strings $x=01234$ and $y=15244$ have $d_H(x,y)=2$, as they differ in two positions.

\subsection{Jaccard Distance} \label{subsec:jaccard-distance}
The \textit{Jaccard Distance} is a popular metric to measure the similarity of sets.
\begin{definition}
	The Jaccard Similarity Coefficient $J(A,B)$ of two sets $A$ and $B$ is given by:
	\begin{equation}
		J(A,B) = \frac{|A \cap B |}{| A \cup B |}
	\end{equation}
\end{definition}
\begin{definition}
	The Jaccard Distance $d_J(A,B)$ between two sets $A$ and $B$ is given by:
	\begin{equation}
		d_J(A,B) = 1 - J(A,B)
	\end{equation}
\end{definition}

This is a valid distance measure because:
\begin{itemize}
  \item it is non-negative: $d_J(A,B) \geq 0 \quad \forall A, B$
  \item it fulfils the identity of indiscernibles: $A = B \Rightarrow d_J(A,B)=0$
  \item it respects the \textit{Triangle Inequality}: $d_J(A,B) \leq d_J(A,C) + d_J(C,B) $
\end{itemize}

For example, the sets $A = \{ 0, 2, 3, 4, 1 \}$ and $B = \{ 1, 3, 5 \}$ have $d_J(A,B) = \frac{2}{6}$, as their intersection is of cardinality 2 and their union of cardinality 6.
